---
title: "Aspen Guide"
output: github_document
date: "2026-08-27"
---

This package provides functions for use in various parts of the data life cycle for data collected for the shared aspen monitoring protocol used by Mojave Desert Network (MOJN) and Upper Columbia Basin Network (UCBN). Currently, functionality exists for creating data package tables for publication, and further functions for QC and visualization are in development. 

### Installation

If you already have a GitHub token for RStudio run the code below to download the aspen package. If you haven’t generated a GitHub token for RStudio, you can follow the guide [here](https://frbcesab.github.io/rsetup/chapters/github-token.html) to set it up and then run the following code to download the aspen package.

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github(repo = "nationalparkservice/mojn-aspen-rpackage")
```


### AGOL Password

In order to access the aspen data, you need AGOL credentials. It is recommended that you use a headless account and that you save the password to your AGOL headless account using the keyring package in R. To add account information to the default keyring follow the steps below. The code should only have to be run once.

It is also recommended that the metadata in your aspen AGOL database is filled out and up to date. Instructions for filling out AGOL metadata can be found [here](https://github.com/nationalparkservice/imd-fetchagol-package. 

```{r, eval = FALSE}
# Run the function below. Change the username field to the username of your headless account and input the password when prompted
keyring::key_set(service = "AGOL", username = "USERNAME", prompt = "Password for headless account: ")

# To check the key was saved run code below, filling in the username to your headless account
keyring::key_get(service = "AGOL", username = "USERNAME")
```

### Functions

Data Publication

- loadAndPackageAspen(): loads aspen data associated with either network from AGOL and wrangles into data package-formatted tables for publication. 

