---
title: "Aspen Quality Control"
format: 
  html:
    embed-resources: true
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(aspen)
```


##### Check Trees
Check that there is at least one tree for every observation entry and flag entries with more than 250 trees

```{r}
if(nrow(aspen:::missingTreeQC()) == 0){
  cat('No errors')
} else{
  rmarkdown::paged_table(aspen:::missingTreeQC())
}
```


##### Check for Missing Species
Check that no entries in the observation table have a SpeciesCode that is NA or missing

```{r}
if(nrow(aspen:::treeSpeciesQC()) == 0){
  cat('No errors')
} else{
  rmarkdown::paged_table(aspen:::treeSpeciesQC())
}
```


##### Check Unknown Trees
Check that unknown species entries do not have any live trees

```{r}
if(nrow(aspen:::unknownSpeciesQC()) == 0){
  cat('No errors')
} else{
  rmarkdown::paged_table(aspen:::unknownSpeciesQC())
}
```


##### Check for Aspens
Check that all of the plots have at least one aspen each time they are visited

```{r}
if(nrow(aspen:::checkAspenQC()) == 0){
  cat('No errors')
} else{
  rmarkdown::paged_table(aspen:::checkAspenQC())
}
```
